Penalized linear mixed models for structured genetic data

### 2021-03-19 main.tex
5.1 (figure still to to), 6

Selection accuracy was evaluated using precision. \anna{All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.} Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs. PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model. For the most part, PC-lasso outperforms LMM-lasso in this setting as well. None of the method are able to accurately select more than about 100 SNPs.  However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first \anna{38} SNPs selected were all false positives.

Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}. Figure~\ref{fig:gw_mse} displays this value for each method and number of selected SNPs. In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.

@@

\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.} \anna{agreed}
The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. However, PC-lasso outperformed LMM-lasso in this case study, in contrast to most of the simulation experiments. This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure. Thus, its structure is well-suited for PC-lasso. This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.

>>>>

Selection accuracy was evaluated using precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. \anna{Figure~\ref{fig:gw_prc}} displays each method's precision for the corresponding number of selected SNPs. PC-lasso and LMM-lasso consistently outperform the ordinary lasso in selecting causal SNPs, particularly for the first $\approx$ 100 SNPs each method allows into the model. For the most part, PC-lasso outperforms LMM-lasso in this setting as well. None of the method are able to accurately select more than about 100 SNPs.  However, because the ordinary lasso does not account for environmental confounding, it cannot reliable select \emph{any} SNPs: the first 38 SNPs selected were all false positives.

Estimation accuracy was assessed via MSE, as defined in Section~\ref{sec:sim-mse}. \anna{Figure~\ref{fig:gw_mse}} displays this value for each method and number of selected SNPs. In accordance with previous simulation results in Section~\ref{sec:sim-mse}, the ordinary lasso results in biased estimates compared to PC and LMM-lasso. This bias is most dramatic for the first $\approx$ 100 SNPs selected by each method. For any number of selected SNPs, however, both PC-lasso and LMM-lasso produce much more accurate estimates than the ordinary lasso.

@@

The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. However, PC-lasso outperformed LMM-lasso in this case study, in contrast to most of the simulation experiments. This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure. Thus, its structure is well-suited for PC-lasso. This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.
